---
title: "MS practicum final presentation"
author: "Zacharia"
date: "6/10/2019"
output: html_document
---

```{r}
library(tibble)
library(dplyr)
library(caret)
library(pROC)
library(MASS)
library(ggplot2)
library(gridExtra)
library(devtools)
library(ggfortify)
library(glmnet)
library(tidyr)
library(naniar)
library(ggsci)

```
#Clean up the data
```{r}
setwd("~/Documents/Research projects/Path project (MS thesis)")

library(readxl)

Main_Data_pathThesis_processed <- read_excel("Main Data_pathThesis_processed copy.xlsx")

head(Main_Data_pathThesis_processed)

rownames(Main_Data_pathThesis_processed)  <- Main_Data_pathThesis_processed$`Pathology Case`

misdiagnosis_predictor <- (Main_Data_pathThesis_processed)

#misdiagnosis_predictor$Pathology Case <- NULL 

dim(misdiagnosis_predictor)

str(misdiagnosis_predictor)
```

#rename columns
```{r}
colnames(misdiagnosis_predictor)[22] <- "other_tissue"

colnames(misdiagnosis_predictor)[23] <- "same_tissue_diagnosis"

colnames(misdiagnosis_predictor)[24] <- "same_diagnosis"

colnames(misdiagnosis_predictor)
```
#remove variables with a lot of missing data and little variablity
```{r}

path_data <- transform(misdiagnosis_predictor,
                     robotic = as.factor(robotic),
                     other_tissue = as.factor(other_tissue),
                    same_diagnosis = as.factor(same_diagnosis),
                    same_tissue_diagnosis = as.factor(same_tissue_diagnosis),
                    tissue_type = as.factor(tissue_type),
                    FS_Speciality = as.factor(FS_Speciality),
                    FS_Pathologist = as.factor(FS_Pathologist),
                    Surgeon = as.factor(Surgeon),
                    Final_pathologist = as.factor(Final_pathologist),
                    Final_speciality = as.factor(Final_speciality),
                    FS_diagnosis = as.factor(FS_diagnosis),
                    Final_diagnosis = as.factor(Final_diagnosis),
                    Origin_tissue = as.factor(Origin_tissue))

#There are many missing variables in the weight column, and the operations are too complex to use in the algorithm.

path_data$Weight <- NULL
path_data$operation <- NULL

#name misdiagnosis levels
levels(path_data$same_diagnosis)[1] <- "Misdiagnosed"
levels(path_data$same_diagnosis)[2] <- "Correct diagnosis"

levels(path_data$robotic)[1] <- "Traditional_Laproscopic"
levels(path_data$robotic)[2] <- "Robotic"


str(path_data)
```

#Divide the data in screening predictor and diagnosis predictor


```{r}
# add a new column

path_data$age_tissue <- (path_data$result_reported - path_data$start_of_operation)

path_data <- path_data[c(1,2,3,4,5,6,23,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22)]

path_data[13,18] <- 42
path_data$Pathology.Case <- NULL

summary(path_data)


# library(Hmisc)
# decribed <- describe(path_data)
# 
# library(summarytools)
# dfSummary(path_data)
```
```{r}
#output the summary into a DF

do.call(cbind, lapply(path_data, summary))

hmsic_df <- data.frame(unclass(Hmisc::describe(path_data)), check.names = FALSE, stringsAsFactors = FALSE)

sumamry_df <- data.frame(unclass(summary(path_data)), check.names = FALSE, stringsAsFactors = FALSE)

write.csv( sumamry_df,"~/Documents/Research projects/Path project (MS thesis)/sumamry_df.csv", row.names = TRUE)

library("psych")
psych_df <- psych::describeBy(path_data, path_data$same_diagnosis)
psych_df_mis <- psych_df$Misdiagnosed
write.csv(psych_df_mis,"~/Documents/Research projects/Path project (MS thesis)/psych_df_mis.csv", row.names = TRUE)
psych_df_correct <- psych_df$`Correct diagnosis`
write.csv(psych_df_correct,"~/Documents/Research projects/Path project (MS thesis)/psych_df_correct.csv", row.names = TRUE)

library(skimr)
skim(path_data)
```
#removing factors
```{r}
path_data_numeric <- path_data[,c(2:7,15,16,18,23)]
summary(path_data_numeric)
```
#Finding correlations
```{r}
as.data.frame(path_data_numeric)

cor(path_data_numeric[,-10])

```
```{r}


library(reshape2)
cormat <- round(cor(path_data_numeric[,-10]),2)
head(cormat)
melted_cormat <- melt(cormat)
head(melted_cormat)

ggplot(data = melted_cormat, aes(Var1,Var2, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
```
#Regularization
```{r}
require(data.table)
require(DT)

fit = glmnet(as.matrix(path_data_numeric[,-c(10)]),as.matrix(misdiagnosis_predictor$same_diagnosis), family="gaussian",alpha=1)

plot(fit,label = TRUE)

print(fit)

coef(fit,s=0.1)

cvfit = cv.glmnet(as.matrix(path_data_numeric[,-10]), as.matrix(misdiagnosis_predictor$same_diagnosis))

plot(cvfit)

require(plotly)
```

```{r}

#getting an aggregate of all varaables by diagnosis type

Aggregated_values_all <- aCRM::Aggregate(x=path_data, by=path_data$same_diagnosis)
write.csv(Aggregated_values_all,"~/Documents/Research projects/Path project (MS thesis)/Aggregated_values_all.csv")

```




```{r}
#t-test for all time related factors

ttest_start_time <- t.test(start_of_operation~ same_diagnosis, path_data)
ttest_date <- t.test(Date_of_operation~ same_diagnosis, path_data)
ttest_result_reported <- t.test(result_reported ~ same_diagnosis, path_data)
ttest_age_tissue <- t.test(age_tissue~ same_diagnosis, path_data)
ttest_age <- t.test(Age~ same_diagnosis, path_data)

ttest_start_time 
ttest_date
ttest_result_reported
ttest_age_tissue 
ttest_age
```


```{r}
#assessing robotic hysterectomy's impact on outcome
robotic_df <- as.data.frame(cbind(Aggregated_values_all$robotic_Traditional_Laproscopic_sum, Aggregated_values_all$robotic_Robotic_sum))
row.names(robotic_df) <- Aggregated_values_all$ID

colnames(robotic_df)[1] <- "laproscopic"
colnames(robotic_df)[2] <- "robotic"

chisq.test(robotic_df)


h <- ggplot(path_data, aes(robotic,same_diagnosis))+
  geom_jitter()+ 
  ylab("F/S diagnosis outcome")

h
```

```{r}
review_complexity <- ggplot(path_data, aes(volume, Review_time)) + geom_point(aes(color = same_diagnosis)) + 
  scale_x_continuous("Volume of tissue (cm3)",limits = c(0, 20))+
  scale_y_continuous("Time tissue reviewed (min)")+
  theme_bw() + labs(title="A") +
  theme(legend.position = "none")

ggsave("~/Documents/Research projects/Path project (MS thesis)/extent_review.jpeg" )

review_complexity
```
```{r}

ttest_volume <- t.test(volume~ same_diagnosis, path_data)
ttest_volume
path_data$age_tissue <- as.numeric(path_data$age_tissue) 

op_complexity <- ggplot(path_data, aes(operation_complexity, age_tissue)) + geom_point(aes(color = same_diagnosis)) + 
  scale_x_continuous("Operation Complexity")+
  scale_y_continuous("Time of Operation (min)")+
  theme_bw() + labs(title="B") +
  theme(legend.title = element_blank())

ggsave("~/Documents/Research projects/Path project (MS thesis)/extent_operation.jpeg" )


```
```{r}
grid.arrange(review_complexity, op_complexity, nrow = 1, layout_matrix = rbind(c(1, 1, 2, 2, 2)))

```

```{r}
# ggplot(path_data, aes(same_diagnosis, fill = FS_Speciality)) + geom_bar()+
# labs(title = "F/S Pathologist speciality by Diagnosis", x = "F/S Diagnosis", y = "Number of Cases")+
#   scale_fill_igv()+
#   theme_minimal()

p2 = ggplot(subset(path_data), 
       aes(same_diagnosis, fill = FS_Speciality)) +
  geom_histogram(colour = "black", binwidth = 1, position = "dodge",  stat = "count") +
  labs(title = "F/S Pathologist speciality by Diagnosis", x = "F/S Diagnosis", y = "Number of Cases")+
  theme_bw()+ scale_fill_igv()+
  theme_minimal()
p2

```
```{r}


# tissue_type_plot<- ggplot(path_data, aes(same_diagnosis, fill = tissue_type)) + geom_bar()+
# labs(title = "Tissue type by Diagnosis", x = "F/S Diagnosis", y = "Number of Cases") +
#    theme( panel.grid.minor = element_blank(),
#         panel.background = element_blank(), axis.line = element_line(colour = "black")) 

tissue_type_plot <- ggplot(subset(path_data), 
       aes(same_diagnosis, fill = tissue_type)) +
  geom_histogram(colour = "black", binwidth = 1, position = "dodge",  stat = "count") +
  labs(title = "Tissue type by Diagnosis", x = "F/S Diagnosis", y = "Number of Cases")+
  theme_bw()+ scale_fill_igv()+
  theme_minimal()


```
```{r reviewing time sample}
means_Review_time <- aggregate( Review_time ~ same_diagnosis, path_data, mean)

t.test(Review_time~same_diagnosis, path_data)

review_sample_plot <- ggplot(path_data,
                             aes(same_diagnosis,Review_time, group = path_data$same_diagnosis)) +
  geom_boxplot(fill = "light blue")+
  scale_y_continuous("Time taken to review sample (min)", breaks= seq(0,3000, by=50), limits = c(0, 120))+
  labs(title = "Time taken to review sample by F/S diagnosis", x = "F/S Diagnosis",
       color = "Discrepent or Correct" ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  geom_jitter(shape = 3, width = 0.2) +
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=2,show_guide = FALSE)

review_sample_plot
```
```{r volume sample}
means <- aggregate(volume ~ same_diagnosis, path_data, mean)

t.test(volume~same_diagnosis, path_data)

volume_plot <- ggplot(path_data,
                             aes(same_diagnosis, volume, group = path_data$same_diagnosis)) +
  geom_boxplot(fill = "light blue") +
  scale_y_continuous("Volume of tissue sent for review (cm3)", breaks= seq(0,3000, by=50), limits = c(0, 120)) +
  labs(title = "A", x = "F/S Diagnosis",
       color = "Discrepent or Correct" ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  geom_jitter(shape = 3, width = 0.2) +
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=2,show_guide = FALSE)
#                 + 
#   geom_text(data = means, aes(means$volume + 0.08, labels = means$volume))
#   
# boxplot(volume ~ same_diagnosis, path_data, ylim=c(0, 1000))
# points(1:2, means$volume, col = "red")
# text(1:2, means$volume + 0.08, labels = means$volume)


volume_plot
```

```{r reviewing sample}
review_sample_plot <- ggplot(path_data,
                             aes(same_diagnosis,Review_time, group = path_data$same_diagnosis)) +
  geom_boxplot(fill = "light blue")+
  scale_y_continuous("Time taken to review sample (min)", breaks= seq(0,3000, by=50), limits = c(0, 120))+
  labs(title = "B", x = "F/S Diagnosis",
       color = "Discrepent or Correct" ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
geom_jitter(shape = 3, width = 0.2)
review_sample_plot 

```
```{r age compared}
age_compared_plot <- ggplot(path_data,
                             aes(same_diagnosis,Age, group = path_data$same_diagnosis)) +
  geom_boxplot(fill = "light blue")+
  scale_y_continuous("Age of Patients (years)", breaks= seq(0,3000, by=50), limits = c(0, 120))+
  labs(title = "C", x = "F/S Diagnosis",
       color = "Discrepent or Correct" ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=2,show_guide = FALSE) +
  geom_jitter(shape = 3, width = 0.2)

age_compared_plot

```

```{r boxplot array}

grid.arrange(volume_plot, review_sample_plot, age_compared_plot, nrow = 1, layout_matrix = rbind(c(1, 2, 3 )))

```
```{r}
library(corrgram)

corrgram(path_data, order=NULL, panel=panel.shade, text.panel=panel.txt,
           main="Correlogram") 


```

```{r}
plot(round(prop.table(table(path_data$same_diagnosis)),2))

#Variables are removed since they are shown to be statistically insignificant
#A new machine learning data frame will be made to show the most statically significant variables

ML_data <- path_data[,c(6:20, 23)]

    train_size <- floor(0.75 * nrow(ML_data))
set.seed(543)
train_pos <- sample(seq_len(nrow(ML_data)), size = train_size)

train_classification <- ML_data[train_pos, ]
test_classification <- ML_data[-train_pos, ]

dim(train_classification)
dim(test_classification)

summary(train_classification)

```

```{r}
library(Boruta) 

set.seed(1407)
randomForest::importance
boruta_analysis = Boruta(same_diagnosis ~ ., data=ML_data, maxRuns=500)
library(mlbench)
plot(boruta_analysis)
as.data.frame(boruta_analysis$finalDecision)


```
#PCA
```{r}

path_data_pca <- prcomp(path_data_numeric[,-c(1,10)],center= TRUE, scale. = TRUE)

summary(path_data_pca)

str(path_data_pca)
```
```{r}
library(devtools)

library(ggbiplot)

#diagnosis_groups <- cut(path_data$same_diagnosis) 

ggbiplot(path_data_pca, obs.scale = 1, groups=path_data$same_diagnosis)

```
```{r}

#PCA by age

age_groups <- cut(path_data$Age, breaks = c(-Inf, 25, 45, 66, Inf), labels = c("Under 25 years", "25-44 years", "45-65 years", "Older than 65"), right = FALSE)

ggbiplot(path_data_pca, ellipse=TRUE, obs.scale = 1, groups=age_groups) + ggtitle("PCA of continous variables grouped by age")
#obs.scale = 1, var.scale = 10,

ggbiplot(path_data_pca, ellipse=TRUE, groups=age_groups, choices = c(3,4)) + ggtitle("3rd and 4th PCA of continous variables grouped by age")
```
```{r}
#PCA by operation complexity

complexity <- cut(path_data$operation_complexity, breaks = c(-Inf, 5, 10, 15, Inf), labels = c("Not extensive", "Average", "Extensive", "Very Extensive"), right = FALSE)

ggbiplot(path_data_pca, ellipse=TRUE, obs.scale = 1,  groups=complexity) + ggtitle("PCA of continous variables grouped by complexity of operation ")


ggbiplot(path_data_pca, ellipse=TRUE, groups=complexity, choices = c(3,4)) + ggtitle("PCA of continous variables in CV data using the 3rd and 4th component")

```
```{r}
#MCA for catagorical variables
library("FactoMineR")
library("factoextra")

path_catagorical <- path_data[,c(8:14,16,17,19,23)]

path_catagorical$operation_complexity <- as.factor(path_catagorical$operation_complexity)

head(path_catagorical)

for (i in 1:8) {
  plot(path_catagorical[,i], main=colnames(path_catagorical)[i],
       ylab = "Count", col="steelblue", las = 2)
}

res.mca <- MCA(path_catagorical, graph = FALSE)
print(res.mca)

eig.val <- get_eigenvalue(res.mca)
fviz_screeplot(res.mca, addlabels = TRUE, ylim = c(0, 45))

var <- get_mca_var(res.mca)
var
```
#Feature selection
##Recursively Filtered data
```{r recurisve filtering}
#set.seed(134)

#winequality_recursive <- path_data
#levels(winequality_recursive$same_diagnosis)[1] <- "1"
#levels(winequality_recursive$same_diagnosis)[2] <- "2"

#svmProfile <- rfe(winequality_recursive[,1:ncol(winequality_recursive)-1],
                 #winequality_recursive$same_diagnosis,
               #   sizes = c(2, 4, 5, 9),
              #    rfeControl = rfeControl(functions = caretFuncs,number = 2),
             #     method = "svmRadial")

#svmProfile
#svmProfile$variables
```
```{r}
#str(winequality_recursive)
```
##Train data GlM
```{r}
library("e1071")
ctrl <- trainControl(method = "repeatedcv", repeats = 20, classProbs = T, savePredictions = T)

levels(train_classification$same_diagnosis) <- make.names(levels(factor(train_classification$same_diagnosis)))

suppressWarnings(path_data_logreg_train <- train(same_diagnosis~., data=train_classification[,-1], method = "glm", family= "binomial", trControl = ctrl))

summary(path_data_logreg_train)
```
```{r}
path_data_logreg_train
```
```{r}
plot(x = roc(predictor = path_data_logreg_train$pred$Misdiagnosed, response = path_data_logreg_train$pred$obs)$specificities, y = roc(predictor = path_data_logreg_train$pred$Misdiagnosed, response = path_data_logreg_train$pred$obs)$sensitivities, col= "blue", xlim = c(1, 0), type ="l", ylab = "Sensitivity", xlab = "Specificity")

legend("bottomright", legend = paste("Misdiagnosed vs. Correctly diagnosed F/S | ", roc(predictor = path_data_logreg_train$pred$Correct.diagnosis, response = path_data_logreg_train$pred$obs)$auc, sep = ""), col = c("blue"), fill = c("blue"))

```
```{r}
#then predict with logistic regression
path_data_predict_class <- predict(path_data_logreg_train, newdata = test_classification)

#confusion matrix
confusionMatrix(path_data_predict_class, reference = test_classification$same_diagnosis)
```

Naive bayes
```{r}
#path_data_numeric[,-c(10,11)] <- NULL

#path_data_numeric <- rbind(path_data_numeric,path_data[,22])

train_size_numeric <- floor(0.75 * nrow(path_data_numeric))
set.seed(543)
train_pos_numeric <- sample(seq_len(nrow(path_data_numeric)), size = train_size)

train_classification_numeric <- path_data_numeric[train_pos, ]
test_classification_numeric <- path_data_numeric[-train_pos, ]

dim(train_classification_numeric)
dim(test_classification_numeric)

summary(train_classification_numeric)

levels(train_classification_numeric$same_diagnosis) <- make.names(levels(factor(train_classification_numeric$same_diagnosis)))

library(naivebayes)
ctrl_bayes <- trainControl(method ="repeatedcv", repeats = 15, classProbs = T, savePredictions = T)

cv_naive_bayes <- train(same_diagnosis~ ., data = train_classification_numeric, trControl = ctrl_bayes, method = 'naive_bayes', na.action=na.exclude)
```
```{r}

 

cv_naive_bayes
```
Naive Bayes ROC curve
```{r}
roc_data <- roc(predictor = cv_naive_bayes$pred$Correct.diagnosis, response = cv_naive_bayes$pred$obs, levels = c("Misdiagnosed", "Correct diagnosis"))$auc

plot(x = roc(predictor = cv_naive_bayes$pred$No.cancer, response = cv_naive_bayes$pred$obs, levels= c("Misdiagnosed", "Correct diagnosis"))$specificities, y = roc(predictor = cv_naive_bayes$pred$No.cancer, response = cv_naive_bayes$pred$obs, levels= c("Misdiagnosed", "Correct diagnosis"))$sensitivities, col= "blue", xlim = c(1, 0), type ="l", ylab = "Sensitivity", xlab = "Specificity")

legend("bottomright", legend = c(paste("Misdiagnosed v Correctly diagnosed --", roc_data, sep = "")), fill =c("blue","red"),  col = c("blue","red"))

```
Prediction and confusion matrix
```{r}
naive_bayes_pred <- predict(cv_naive_bayes, newdata = test_classification[-7])
confusionMatrix (cv_naive_bayes, reference = test_classification$same_diagnosis)
```
##Trees
```{r}
library(randomForest)
library(tree)
library(caret)
library(pROC)
library(gbm)
library(ROCR)

set.seed(543)
regression_tree <- tree(same_diagnosis ~ ., data = train_classification_numeric)

plot(regression_tree)
text(regression_tree, cex=0.75)

library(party)
irisct <- ctree(same_diagnosis ~ ., data = path_data_numeric)
plot(irisct)

```
```{r}

summary(regression_tree)
```


Test this tree on the test set
```{r}
classification_test <- predict(regression_tree, newdata = test_classification_numeric, type = "class")

confusionMatrix(classification_test, reference = test_classification_numeric$same_diagnosis)
```

Fit the tree using cross validation. Use FUN = prune.misclass to indicate we want to classification error to guide cross val and pruning. 
```{r}
fit_classification_tree <- cv.tree(classification_test,FUN=prune.misclass, K = 13)

fit_classification_tree
```
Shows trees with 19 or 1 nodes do the best

Now prune the tree 
```{r}
prune_classification_tree=prune.misclass(regression_tree, best=11)

plot(prune_classification_tree)
text(prune_classification_tree,cex=0.45)
```

Test the pruned tree on the test set
```{r}
classification_test_pruned <- predict(prune_classification_tree, newdata = test_classification, type = "class")

confusionMatrix(classification_test_pruned, reference = test_classification$same_diagnosis)
```
#Random Forest

```{r}
na.omit(train_classification)

isError <- is.na(train_classification) 


RF_classification <- randomForest(same_diagnosis~., data= train_classification_numeric,  importance = TRUE, oob.times = 15, confusion = TRUE)

RF_classification

predictions_RF <- predict(RF_classification, newdata = test_classification_numeric)

confusionMatrix(predictions_RF, reference = test_classification_numeric$same_diagnosis)
```

#Support Vector Machine

Train linear SVM
```{r}
set.seed(30495)
ctrl <- trainControl(method = "repeatedcv", repeats = 5,classProbs = T, savePredictions = T)
svm <- train(same_diagnosis~ ., data = train_classification, method = "svmLinear", tuneLength = 10, trControl = ctrl)

svm
```

```{r}
roc(predictor = svm$pred$Misdiagnosed, response = svm$pred$obs)$auc
```

Visualize ROC curve for SVM
```{r}
plot(x = roc(predictor = svm$pred$Misdiagnosed, response = svm$pred$obs)$specificities, y = roc(predictor = svm$pred$Misdiagnosed, response = svm$pred$obs)$sensitivities, col= "blue", xlim = c(1, 0), type ="l", ylab = "Sensitivity", xlab = "Specificity")

legend("bottomright", legend = paste("Misdiagnosed vs. correctly diagnosed F/S | ", roc(predictor = svm$pred$Misdiagnosed, response = svm$pred$obs)$auc, sep = ""), col = c("blue"), fill = c("blue"))
```
lasso regression
```{r}
CARET.TRAIN.CTRL <- trainControl(method="repeatedcv",
                                 number=5,
                                 repeats=5,
                                 verboseIter=FALSE)

# test out Ridge regression model

lambdas <- seq(1,0,-0.001)

Y <- path_data$same_diagnosis 


path_data_numeric$same_diagnosis <- NULL
# train model
set.seed(123)  # for reproducibility
model_ridge <- cv.glmnet(x=as.matrix(path_data_numeric),y=Y,alpha=1,family="binomial")

model_ridge$lambda.1se
plot(model_ridge)

```


```{r}
#summary(path_data_logreg)$coef

```

```{r}
path_data_probability <- predict(path_data_logreg_train, type="response")



Log_regression_predictions <- rep("No screening needed",668)
Log_regression_predictions [path_data_probability >.5] <-  "Should have screened"

table(Log_regression_predictions, path_data$same_diagnosis)

```
```

```{r}
